An efficient acyclic contact planner for multiped robots

### 2016-01-18 body/overview.tex
nico + glossary

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed in a sequential fashion.
The guarantees provided by the \textit{geometrically feasible} guide path allow
us to avoid the combinatorial by generating the contacts in a straightforward manner.

%
\begin{figure*}
  \centering
  \begin{overpic}[width=0.8\linewidth]{figures/contact_gen}
		\put (1,1) {a} 
		\put (22,1) {b} 
		\put (42,1) {c} 
		\put (62,1) {d} 
		\put (83,1) {e} 
	\end{overpic}
  \caption{Generation of a contact configuration for the right leg of HRP-2. a: Selection of reachable obstacles. b: Entries of the limb samples database (with $N = 4$). c: With a proximity query on the octree database, configurations too far from obstacles are eliminated. d: The best candidate according to a user-defined heuristic $h$ is chosen. e: The final contact is achieved using inverse kinematics.}
  \label{fig:contact_gen}
\end{figure*}
\subsection{Computation of a guide path --- \Pa}
We first consider the problem of planning a relevant guide path. The objective is to compute a path of root placements that will allow contact creation. A desired requirement is to preserve the completeness of the planner: it should be able to explore any possible guide path, but at the same time, any computed guide path must be \textit{truly feasible}, i.e. must lead to a valid sequence of contacts.

More precisely, a root placement is \textit{truly feasible} if and only if there exists at least one joint configuration such that the robot satisfies three conditions:
\begin{description}
\item[C1] is not in collision with the environment, 
\item[C2] has at least one end-effector in contact with the environment, 
\item[C3] is in static equilibrium. 
\end{description}
C1 and C2 define \textit{geometrical feasibility}, and can be efficiently approximated even without computing the contact points. C3 defines \textit{equilibrium feasibility}. C3 is difficult to approximate and requires the computation of the contact points, as well as the resolution of a linear program to compute the contact forces~\citep{Prete2016}. 
However our intuition suggests us that, for cluttered environments, verifying C3 is not critical for $\mathcal{P}_1$, meaning that, most of the times, root placements satisfying C1 and C2 would satisfy also C3. For this reason we neglect C3 during the resolution of \Pa, and we worry about it only when solving \Pb.

An intuitive description of root placements satisfying C1 and C2 is ``close, but not too close'': close, because a contact surface must be partially included in the reachable workspace of the robot (represented for the right arm in Figure~\ref{fig:contact_gen}--a); not too close, because the robot must avoid collision. We define formally $C_{reach}$, the set of root placements satisfying C1 and C2, in which we compute a guide path with a sampling-based planner, the Reachability-Based RRT---RB-RRT--- (Figure~\ref{fig:framework}---\Pa). Planning in $C_{reach}$ boils down to planning in SE(3), which has an acceptable practical complexity.
%
Details are presented in Section~\ref{rbprm}.

@@

The second stage extends the guide path into a sequence of contact configurations (Figure~\ref{fig:framework}--\Pb). In this
step the input guide path is \textit{geometrically feasible}.
With a fixed root position, the dimensionality of the 
contact generation problem is reduced to the number of degrees of freedom of the considered limb. In other words, we consider each limb as a manipulator attached to the root, and select the most relevant contact from a database of precomputed configurations, which is independent from the environment (Figure~\ref{fig:contact_gen}).

From a given start configuration, the planner proceeds in an iterative fashion along the discretized path: given a new root placement, an inverse-kinematics solver 
is used to maintain the contacts that were previously existing. Often, these contacts cannot be maintained, because of joint limits or collisions.
In such a case the contacts are broken. Conversely, new contacts are created to ensure the static equilibrium of the robot.
The algorithm is designed so that only one contact can be created or broken between two successive configurations. This ensures that the interpolation between two configurations is achievable.
Details are presented in Section~\ref{sec:contact}.

To select the ``most relevant'' contact configurations, user-defined heuristics can be chosen. In Section~\ref{sec:heuristics} we present extensively two such heuristics, which have been used in our experiments.

@@

\begin{figure}
  \centering
  \begin{overpic}[width=0.8\linewidth]{figures/character}
    \put (7,29) {\small{$\in \mathbf{q}^1$}}
    \put (7,25) {\small{effector}}
  \end{overpic}
  \caption{
    Left: Robot in a rest configuration. The right arm is denoted as the limb $R^1$. Each colored dot represents a degree of freedom around an axis. Right: Volumes of the robot. The red geometry denotes $W^0$ and must remain collision-free. The green spheres are the reachable workspace of each limb, the  $W^k$.}
  \label{fig:character}
\end{figure}

\medskip
\textbf{A robot} is a kinematic chain $R$, comprising \mbox{$n + 6$} degrees of freedom (DOFs).
$R$ is composed of $l$ limbs $R^k, 1 \leq k \leq l$, attached to a root.
It is described by a configuration $\mathbf{q} \in SE(3) \times \mathbb{R}^n$.
We define some relevant projections of $\mathbf{q}$:
\begin{itemize}
	\item $\mathbf{q}^k$ denotes the configuration (a vector of joint values) of the limb $R^k$ (Figure~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ denotes the vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in SE(3)$ denotes the position and orientation of the root of the robot $R$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
The volume encompassing the trunk of the robot is denoted $W^0$ (Figure~\ref{fig:character}-right: central cylinder). The reachable workspace of a limb $R^k$ is denoted $W^k$ (Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{jl}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\},
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0$ being the null displacement, and  $C^k_{jl}$ is the space
of admissible limb joint configurations. We also define $W = \bigcup_{k=1}^{l}W^k$, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ translated and rotated by the rigid displacement~$\mathbf{q}^{0}$.

>>>>

%~ \begin{table*}
%~ \centering
%~ \begin{tabular}{ c | >{\arraybackslash}m{200pt} | c}
   %~ Robot configuration set &  \centering Definition: $\mathbf{q} \in$ set $\Rightarrow$?  & Root configuration set\\
 %~ \hline
 %~ \hline
   %~ $C_{contact}$ & $\mathbf{q}$ collision free; at least one end effector in contact. &  $\mathbf{q}^{0} \in C_{contact}^0$ is ``contact feasible''.  \\
 %~ \hline
   %~ $C_{equil} \subset C_{contact} $ & $\mathbf{q}$  in static equilibrium &  $\mathbf{q}^{0} \in C_{equil}^0$ is ``equilibrium feasible''. \\
 %~ \end{tabular}
%~ \caption{TODO.}
%~ \label{tab:glossary}
%~ \quad
%~ \end{table*}

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed in a sequential fashion: we first plan a guide path with the \textit{reachability condition}, before
extending it into a sequence of contact configurations in static equilibrium.
In this Section we describe this workflow and give some notations used throughout the paper. Some key concepts are introduced,
the definitions of which can be found in the glossary at the end of this paper.

%
\begin{figure*}
  \centering
  \begin{overpic}[width=0.8\linewidth]{figures/contact_gen}
		\put (1,1) {a} 
		\put (22,1) {b} 
		\put (42,1) {c} 
		\put (62,1) {d} 
		\put (83,1) {e} 
	\end{overpic}
  \caption{Generation of a contact configuration for the right leg of HRP-2. (a): Selection of reachable obstacles. (b): Entries of the limb samples database (with $N = 4$). (c): With a proximity query on the octree database, configurations too far from obstacles are eliminated. (d): The best candidate according to a user-defined heuristic $h$ is chosen. (e): The final contact is achieved using inverse kinematics.}
  \label{fig:contact_gen}
\end{figure*}

\subsection{Computation of a guide path --- \Pa}
We first consider the problem of planning a relevant guide path. The objective is to compute a path of root configurations that will allow contact creation. A desired requirement is to preserve the completeness of the planner: it should be able to explore any valid guide path, but at the same time, any computed guide path must lead to a sequence of contact configurations in static equilibrium.

More precisely, a root configuration is \gls{equilibrium feasible} if and only if there exists at least one joint configuration such that the robot satisfies two conditions:
\begin{description}
\item[C1]: is not in collision with the environment, 
%~ \item[C2] has at least one end-effector in contact with the environment, 
\item[C2]: is in static equilibrium. 
\end{description}
A requirement for C2 is the condition C3: at least one end-effector must be in contact with the environment.
C1 and C3 define \glslink{contact feasible}{\textit{contact feasibility}}, and can be efficiently approximated even without computing the contact points. C2 defines \glslink{equilibrium feasible}{\textit{equilibrium feasibility}}. C2 is difficult to approximate and requires the computation of the contact points, as well as the resolution of a linear program to compute the contact forces~\citep{Prete2016}.

We make the assumption that for the \textit{\gls{cluttered} contact planning problem} (that we define as admitting a solution where every configuration contains at least one contact
is \gls{quasi-flat}) verifying C2 is not critical for $\mathcal{P}_1$.
This means that root configurations satisfying C1 and C3 also satisfy C2. For this reason we neglect C2 during the resolution of \Pa, and we worry about it only when solving \Pb.

We define $C_{Contact}^0$ as the set of \gls{contact feasible} root configurations (that can satisfy C1 and C2).
Planning in $C_{Contact}^0$ boils down to planning in $\mathbb{R}^r$, which has an acceptable practical complexity.

An intuitive description of \gls{contact feasible} configurations is ``close, but not too close'': close, because a contact surface must be partially included in the reachable workspace of the robot (represented for the right leg in Figure~\ref{fig:contact_gen}--a); not too close, because the robot must avoid collision.
We approximate $C_{Contact}^0$ with a set $C_{reach}^0$, defined by the \textit{reachability condition}, a geometrical criterion based on this intuitive description.
In Section~\ref{rbprm} we detail how  $C_{reach}^0$ is sampled to compute a guide path with a dedicated planner, the Reachability-Based RRT---RB-RRT--- (Figure~\ref{fig:framework}---\Pa).

@@

The second stage extends the guide path into a sequence of contact configurations (Figure~\ref{fig:framework}--\Pb).
With a fixed root position, the dimensionality of the 
contact generation problem is reduced to the number of degrees of freedom of the considered limb. In other words, we consider each limb as a manipulator attached to the root, and select the most relevant contact from a database of precomputed configurations, which is independent from the environment (Figure~\ref{fig:contact_gen}).

From a given start configuration, the planner proceeds in an iterative fashion along the discretized path: given a new root configuration, an inverse-kinematics solver 
is used to maintain the contacts that were previously existing. Possibly, some of these contacts cannot be maintained (because of joint limits or collisions).
Then the contacts are broken. Conversely, new contacts are created to ensure the static equilibrium of the robot.
The algorithm is designed so that only one contact can be created or broken between two successive configurations. This ensures that the interpolation between two configurations is achievable.
Details are presented in Section~\ref{sec:contact}.

There is a combinatorial issue in the choice of which contact to create,
and where to create it. This complexity is avoided by selecting the ``most relevant'' contact configurations with user-defined heuristics. In Section~\ref{sec:heuristics} we present extensively those heuristics that have been used in our experiments.

@@

%~ \begin{figure}
  %~ \centering
  %~ \begin{overpic}[width=0.8\linewidth]{figures/character}
    %~ \put (7,29) {\small{$\in \mathbf{q}^1$}}
    %~ \put (7,25) {\small{effector}}
  %~ \end{overpic}
  %~ \caption{
    %~ Left: Robot in a rest configuration. The right arm is denoted as the limb $R^1$. Each colored dot represents a degree of freedom around an axis. Right: Volumes of the robot. The red geometry denotes $W^0$ and must remain collision-free. The green spheres are the reachable workspace of each limb, the  $W^k$.}
  %~ \label{fig:character}
%~ \end{figure}

\medskip
\textbf{A robot} is a kinematic chain $R$, comprising \mbox{$n + r$} degrees of freedom (DOFs), with $r \geq 6$. $r$ denotes the world coordinates of the root of the robot,
as well as extra DOFs not related to the limbs (For instance, HRP-2 has two extra DOFS in the torso, such that we have $r=8$. 
$R$ is composed of $l$ limbs $R^k, 1 \leq k \leq l$, attached to a root.
It is described by a configuration $\mathbf{q} \in  \mathbb{R}^r \times \mathbb{R}^n$.
We define some relevant projections of $\mathbf{q}$:
\begin{itemize}
	\item $\mathbf{q}^k$ denotes the configuration (a vector of joint values) of the limb $R^k$; % (Figure~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ denotes the vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in \mathbb{R}^r$ denotes the position and orientation of the root of the robot $R$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
The volume encompassing the trunk of the robot is denoted $W^0$
 %~ (Figure~\ref{fig:character}-right: central cylinder). 
 The reachable workspace of a limb $R^k$ is denoted $W^k$ %(Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{jl}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\},
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0$ being the null displacement, and  $C^k_{jl}$ is the space
of admissible limb joint configurations. We also define $W = \bigcup_{k=1}^{l}W^k$, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ translated and rotated by the rigid displacement~$\mathbf{q}^{0}$.

### 2016-01-18 body/rbprm.tex
nico + glossary

We consider the issue of computing a normalized guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow SE(3)$ for the root of a multiped robot, connecting user-defined start and goal configurations. As said in the previous section, the goal is to enforce that any configuration $\mathbf{q}^0$ of the guide is geometrically feasible, i.e. can be mapped to a configuration in contact. We denote by $C_{contact} \subset SE(3) \times \mathbb{R}^n$ the contact sub-manifold of the robot, that is the set of configurations for which the robot is in contact with the environment.
%~ We say that a root placement $\mathbf{q}^{0}$ is \textit{truly feasible} if there exists $\mathbf{q}^{\overline{0}}$ such that 
%~ $$\mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{contact} $$
%~ The set of all truly-feasible root placements is denoted by $C_{reach}$. By extension, a trajectory $\mathbf{q}^0(t)$ is truly feasible if $\forall t \in [0,1], \mathbf{q}^0(t) \in C_{reach}$. 
%~ We define a bijective mapping $\pi$ between $C_{reach}$ and $C_{contact}$:
%~ \begin{equation}
    %~ \pi\colon \quad \mathbf{q}^{0} \in C_{reach} \longrightarrow  \mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{contact} 
    %~ \label{eq:pi}
%~ \end{equation}

We say that a root placement $\mathbf{q}^{0}$ is \textit{geometrically feasible} if there exists at least one bijective mapping $\pi$
such that
\begin{equation}
    \pi\colon \quad \mathbf{q}^{0} \in  SE(3) \longrightarrow  \mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{contact} 
    \label{eq:pi}
\end{equation}

The set of all \textit{geometrically feasible} root placements is denoted by $C_{reach}$. By extension, a path $\mathbf{q}^0(t)$ is \textit{geometrically feasible} if $\forall t \in [0,1], \mathbf{q}^0(t) \in C_{reach}$.

For this stage, and for this stage only, we make the assumption that in cluttered environments, a \textit{geometrically feasible} path can be transformed into a \textit{truly feasible} path.
Under this assumption, for a two-stage acyclic contact planner to be exact and complete, we need the combination of two conditions on a guide path generator: all the generated paths must be \textit{geometrically feasible} (sufficient condition); the guide planner must be complete, i.e. it must be able to discover any \textit{geometrically feasible} path (necessary condition).

In this Section we characterize $C_{reach}$ theoretically with a necessary and a sufficient condition for belonging to the set.
We then describe our implementation of an efficient RRT planner in $C_{reach}$.

@@

\subsection{Conditions for geometrical feasibility}
By default, \textit{geometrical feasibility} implies a constructive demonstration by exhibiting a valid $\pi$. This is the approach chosen by \cite{Bouyarmane2009}. However, computing a valid  $\mathbf{q}^{\overline{0}}$ is costly in practice. In this section we rather define a necessary condition and a sufficient condition for \textit{geometrical feasibility} that do not require this explicit computation.

% and  present the parametrization of a compromise condition used to approximate $C_{reach} \subset SE(3)$, the set of truly feasible configurations. This compromise condition, which we call the reachability condition, is used to compute a relevant guide trajectory using a sampling based motion planner, the Reachability-Based PRM (RB-PRM).

\subsubsection*{Geometrical feasibility: necessary condition}
For a contact to be possible, a volume $O_i \in O$ necessarily intersects with the reachable workspace $W(\mathbf{q}^{0})$ (Figure~\ref{fig:contact_gen}--1). Furthermore, if $\mathbf{q}^{0}$ is  feasible, then the trunk of the robot $W^0(\mathbf{q}^{0})$ is necessarily not colliding  with the environment $O$.

Therefore we can approximate $C_{reach}$ with a set $C_{reach} \subset C_{reach}^1$ with the reachability condition defined as: 
\begin{equation}
C_{reach}^1 = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge W^0(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \wedge & A_{trunk}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
It is straightforward to prove that  $C_{reach} \subset C_{reach}^1$ (by construction of the included set). 
This inclusion is very important: it directly implies that any motion-planning algorithm with a guaranty of completeness in $SE(3) \times \mathbb{R}^n$ is complete  in $C_{reach} \times \mathbb{R}^n$. This is a strong reduction of the search space, which can be directly applied to any existing method.

The condition $C_{reach}^1$ is only necessary which means that one such root placement might not be truly feasible: in practice it is not guaranteed to find a valid sequence of contacts for every guide path in $C_{reach}^1$.

\subsubsection*{Geometrical feasibility: sufficient condition}
%~ In building $C_{reach}^1$, we directly considered the including hull of the root body $W^{0}$, obtaining a necessary condition. 
A trivial sufficient condition for \textit{geometrical feasibility} can be constructed as a variation of $C_{reach}^1$, by replacing $W^0$ with a bounding volume $B^{max}$ encompassing the whole robot in a given pose, except for the effector surfaces to be in contact. We denote by \mbox{$C_{reach}^\infty \subset C_{reach}$} the set of root placements corresponding to the sufficient condition.

In general, the inclusion is strict, which means that we lose the completeness of the two-stage contact planner (i.e. the planner is not able to discover a path inside \mbox{$C_{reach} \setminus C_{reach}^\infty$}). However, the sufficient condition guarantees that any such path leads to a valid sequence of contacts (i.e. $\pi$ is defined).

\subsection{Geometrical feasibility: a compromise reachability condition}
The sufficient condition is not interesting in practice since it leads the solver to lose too many interesting paths. The necessary condition is not perfect either, since the first stage of the planner would stop on a guide that is not \textit{geometrically feasible} in practice. It might be possible to find a shape $B$ that is necessary and sufficient; however, it seems intuitively very unlikely in general. The construction of a shape $W^0 \subset B \subset B^{max}$ leading to a necessary and sufficient condition (or the proof of its nonexistence) is out of the scope of this work.

However, between $W^0$ and $B^{max}$, a trade-off can be found between a necessary and a sufficient condition. We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$. 
%
We then consider the spaces $C_{reach}^s$
 \begin{equation}
C_{reach}^s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \wedge & A_{trunk}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
The higher $s$, the closer this ``reachability condition'' is to being sufficient, and if $s=1$, the planner is complete. The parametrization of $s$ allows to find a trade-off between these two desirable properties.
As an example, Figure~\ref{fig:HyQ_roms} presents the volumes $W$ computed for the HyQ robot.
In Appendix~\ref{app:rom}, we give a generic method to compute these volumes, with the example of HRP-2.
%~ Section~\ref{sec:results} shows that in practice, it is easy to adjust $s$ to keep most of the interesting guides without introducing incorrect guides.

@@

\subsection{Computing the guide path in $C_{reach}^s$}
Once a value of $s$ has been fixed, any sampling-based motion planner can be used to plan a path in $C_{reach}^s$. 
Indeed, contrary to $C_{Contact}$,  $C_{reach}^s$ has a non zero measure in the configuration space $C$. Therefore a standard uniform sampling approach
can work, in spite of a high rejection rate. 
Thus, the only significant change regarding a classical planner is to replace the classical collision checking method with a test of belonging to $C_{reach}^s$ when verifying
that drawn configurations and associated local paths are valid.

However, to improve the sampling efficiency while remaining probabilistically complete, we bias the sampling process to generate near obstacle configurations, similarly to~\cite{Amato98choosinggood}.
First, a configuration is set to a random point on the surface of one randomly chosen obstacle. The root location is then translated and rotated randomly until the reachability condition is satisfied.
Our current implementation of these modifications is based on the Bi-RRT planner provided by the HPP software.

Thanks to this small set of modifications, the high dimensional truly feasible path planning problem is reduced to a standard geometry and collision checking problem, in a much lower dimension (6 for HyQ, 8 for HRP-2 that has joints in the torso). In so doing we make the problem tractable for sampling-based approaches, with extremely efficient computation times.

>>>>

We consider the issue of computing a normalized guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow \mathbb{R}^r$ for the root of a multiped robot, connecting user-defined start and goal configurations (. As said in the previous section, the goal is to enforce that the guide can be discretized into a sequence of \gls{contact feasible} configurations. We denote by \gls{ccontact} the contact sub-manifold of the robot.

We say that a root configuration $\mathbf{q}^{0}$ is \gls{contact feasible} if it belongs to the set:
\begin{equation}
      C_{contact}^0= \{ \mathbf{q}^{0}: \exists \mathbf{q}^{\overline{0}}, \mathbf{q}^{0} \oplus \mathbf{q}^{\overline{0}} \in C_{contact} \}
    \label{eq:pi}
\end{equation}

%~ The set of all \textit{geometrically feasible} root configurations is denoted by $C_{reach}$.
By extension, a path $\mathbf{q}^0(t)$ is \gls{contact feasible} if $\forall t \in [0,1], \mathbf{q}^0(t) \in C_{contact}^0$.

For this stage, and for this stage only, we make the assumption that for \gls{cluttered} problems, a \gls{contact feasible} path can be transformed into a \gls{equilibrium feasible} path.
Under this assumption, for a two-stage acyclic contact planner to be exact and complete, we need the combination of two conditions on a guide path generator: all the generated paths must be \gls{contact feasible} (sufficient condition); the guide planner must be complete, i.e. it must be able to discover any \textit{contact feasible} path (necessary condition).

In this Section we characterize \gls{ccontact0} theoretically with a necessary and a sufficient condition for belonging to the set.
We then approximate \gls{ccontact0} as a space $C_{reach}$, where we can efficiently sample configurations.
We finally describe our implementation of an efficient RRT planner in $C_{reach}$.

@@

\subsection{Conditions for contact feasibility}
By default, \glslink{contact feasible}{\textit{contact feasibility}} implies a constructive demonstration by exhibiting a valid $\mathbf{q}^{\overline{0}}$. This is the approach chosen by \cite{Bouyarmane2009}. However, computing $\mathbf{q}^{\overline{0}}$ is expensive in terms of computation time. In this section we rather define a necessary condition and a sufficient condition for \glslink{contact feasible}{\textit{contact feasibility}} that do not require this explicit computation.

\subsubsection*{Contact feasibility: necessary condition}
For a contact to be possible, a volume $O_i \in O$ necessarily intersects with the reachable workspace $W(\mathbf{q}^{0})$ (Figure~\ref{fig:contact_gen}--1). Furthermore, if $\mathbf{q}^{0}$ is  feasible, then the trunk of the robot $W^0(\mathbf{q}^{0})$ is necessarily not colliding  with the environment $O$.

Therefore we can define an optimistic approximation  $C^0_{Nec} \supset$ \gls{ccontact0} defined as: 
\begin{equation}
C_{Nec}^0 = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } W^0(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \wedge & A_{trunk}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
The inclusion \gls{ccontact0} $\subset C^0_{Nec}$ is straightforward, but it is very important.
Planning in $C^0_{Nec}$ allows a strong reduction of the search space, while not discarding 
any root configuration that could lead to a solution.

The condition defining $C^0_{Nec}$ is only necessary. This means that it might not be possible to extend a guide path planned 
in $C^0_{Nec}$ into a sequence of contact configurations.

\subsubsection*{Contact feasibility: sufficient condition}
%~ In building $C_{reach}^1$, we directly considered the including hull of the root body $W^{0}$, obtaining a necessary condition. 
A trivial sufficient condition for \glslink{contact feasible}{\textit{contact feasibility}} can be constructed as a variation of $C^0_{Nec}$, by replacing $W^0$ with a bounding volume $B^{Suf}$ encompassing the whole robot in a given pose, except for the effector surfaces to be in contact. We denote by \mbox{$C_{Suf}^0 \supset $ \gls{ccontact0}} the set of root configurations corresponding to this sufficient condition.

In general, the inclusion is strict, which means that we lose the completeness of the two-stage contact planner (i.e. the planner is not able to discover a path inside \mbox{$C_{reach} \setminus C_{Suf}^0$}). However, the sufficient condition guarantees that any such path leads to a valid sequence of contacts.

\subsection{Contact feasibility: a compromise \gls{reachc}}
The sufficient condition is not interesting in practice since it leads the solver to lose too many interesting paths. The necessary condition is not perfect either, since the first stage of the planner would stop on a guide that is not \gls{contact feasible} in practice. It might be possible to find an ideal shape $B$ that is necessary and sufficient; however, it seems intuitively very unlikely in general. The construction of a shape $W^0 \subset B \subset B^{Suf}$ leading to a necessary and sufficient condition (or the proof of its nonexistence) is out of the scope of this work.

However, between $W^0$ and $B^{Suf}$, a trade-off can be found between a necessary and a sufficient condition. We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$.
%
We then consider the spaces $C_{s}^0$
 \begin{equation}
C^0_s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \wedge W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{trunk}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
If $s=1$, then $W^0_s$ = $W^0$, such that $C_1^0$ = $C_{Nec}$. We thus define $s \in \mathbb{R}^+$, since values inferior to $1$ can only introduce more errors.
The higher $s$, the closer this \gls{reachc} is to being sufficient, and if $s=1$, the search is complete (in the sense that all \gls{contact feasible} configurations can be found). The parametrization of $s$ defines a trade-off between these two desirable properties. $s$ can be chosen by hand, or automatically as explained in section TODO.
In the rest of the paper, we consider that an appropriate value $s^*$ of $s$ is chosen, and define \gls{creach0} $= C^0_{s^*}$.

As an example, Figure~\ref{fig:HyQ_roms} presents the volumes $W$ computed for the HyQ robot.
In Appendix~\ref{app:rom}, we give a generic method to compute these volumes, with the example of HRP-2.
%~ Section~\ref{sec:results} shows that in practice, it is easy to adjust $s$ to keep most of the interesting guides without introducing incorrect guides.

@@

\subsection{Computing the guide path in $C_{reach}^0$}
Any sampling-based motion planner can be used to plan a path in \gls{creach0}. 
Indeed, contrary to \gls{ccontact}, \gls{creach0} has a non zero measure in the configuration space $C$. Therefore a standard uniform sampling approach
can work, in spite of a high rejection rate. 
Thus, the only significant change regarding a classical planner is to replace the classical collision checking method with a test of belonging to \gls{creach0} when verifying
that drawn configurations and associated local paths are valid.

However, to improve the sampling efficiency while remaining probabilistically complete, we bias the sampling process to generate near obstacle configurations, similarly to~\cite{Amato98choosinggood}.
First, a configuration is set to a random point on the surface of one randomly chosen obstacle. The root location is then translated and rotated randomly until the \gls{reachc} is satisfied.
Our current implementation of these modifications is based on the Bi-RRT planner \citep{770022} provided by the HPP software.

Thanks to this set of modifications, the problem of planning an \gls{contact feasible} path in a high dimensional space is reduced to a standard geometry and collision checking problem, in a much lower dimension (6 for HyQ, 8 for HRP-2 that has joints in the torso). In so doing we make the problem tractable for sampling-based approaches, with \gls{interactive} computation times.
